---
title: "p8105_hw1_rd3096"
author: "Ruicong(Timothy) Du"
date: "2024-09-15"
output: github_document
---
#problem 1

##load the penguins dataset
```{r setup}
library(palmerpenguins)
data("penguins", package = "palmerpenguins")
```

##show structure of the dataset for variables and their values
```{r explore dataset}
str(penguins)
```

## description of the dataset
The penguins dataset contains information on penguins from the Palmer Archipelago, Antarctica. The dataset consists of `r nrow(penguins)` rows and `r ncol(penguins)` columns.

The key variables and values are:
*species: The species of penguins (Adelie, Chinstrap, Gentoo).
*island: The island where the penguins were observed (Biscoe, Dream, Torgersen).
*bill_length_mm: The length of penguin bills in millimeters.
*flipper_length_mm: The length of penguin flippers in millimeters.
*body_mass_g: The body mass of penguins in grams.
*sex: The sex of the penguins (male or female).

The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.


