---
title: "p8105_hw1_rd3096"
author: "Ruicong(Timothy) Du"
date: "2024-09-15"
output: github_document
---
#problem 1

##load the penguins dataset
```{r setup, message = FALSE, warning = FALSE}
library(ggplot2)
library(tidyverse)
data("penguins", package = "palmerpenguins")
```
##show structure of the dataset for variables and their values
```{r explore dataset}
str(penguins)
```

## description of the dataset
The penguins dataset contains information on penguins from the Palmer Archipelago, Antarctica. The dataset consists of `r nrow(penguins)` rows and `r ncol(penguins)` columns.

The key variables and values are:
*species: The species of penguins (Adelie, Chinstrap, Gentoo).
*island: The island where the penguins were observed (Biscoe, Dream, Torgersen).
*bill_length_mm: The length of penguin bills in millimeters.
*flipper_length_mm: The length of penguin flippers in millimeters.
*body_mass_g: The body mass of penguins in grams.
*sex: The sex of the penguins (male or female).

The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.

##Make a scatterplot for flipper length vs bill length
```{r scatterplot, warning=FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  labs(
    title = "Flipper Length vs Bill Length by Species",
    x = "Bill Length (mm)",
    y = "Flipper Length (mm)",
    color = "Species"
  ) 
```

#problem 2

##create the data frame
```{r data frame}
set.seed(1)
frame_df=tibble(
  vec_numeric=rnorm(10),
  vec_logical=rnorm(10)>0,
  vec_char=letters[1:10],
  vec_factor=factor(rep(c("level1", "level2", "level3"), length.out = 10))
)

print(frame_df)
```

##take the mean of each variable
```{r mean,warning=FALSE}
mean_numeric=mean(pull(frame_df,vec_numeric))
mean_logical=mean(pull(frame_df,vec_logical))
mean_char=tryCatch(mean(pull(frame_df,vec_char)), error = function(e) e)
mean_factor=tryCatch(mean(pull(frame_df,vec_factor)), error = function(e) e)
```

The mean of the numeric vector is `r mean(frame_df$vec_numeric)`.

The mean of the logical vector is `r mean(frame_df$vec_logical)`.

For the character vector, attempting to calculate the mean results in: `r tryCatch(mean(frame_df$vec_char), error = function(e) "Error: Cannot calculate mean for characters")`.

For the factor vector, attempting to calculate the mean results in: `r tryCatch(mean(frame_df$vec_factor), error = function(e) "Error: Cannot calculate mean for factors")`.


##Explanation for what mean works and what doesnot

1.mean_numeric: This will calculate the mean of vec_numeric, which is a numeric vector. This will work because the values are numeric.

2.mean_logical: This will calculate the mean of vec_logical. Logical vectors are treated as 1 (TRUE) and 0 (FALSE), so the mean will represent the proportion of TRUE values.

3.mean_char: This will fail because the vec_char is a character vector, and r cannot calculate a mean for characters. 

4.mean_factor: This will also fail because factors are categorical variables. and r cannot calculate a mean for characters. 

## Explicit conversion to numeric
```{r chunk_name, echo=TRUE, results='hide'}
as.numeric(pull(frame_df, vec_logical))
as.numeric(pull(frame_df, vec_char))
as.numeric(pull(frame_df,vec_factor))
```

##Explaination for coercion and convertion:

1.Logical vectors are treated as 1 (TRUE) and 0 (FALSE), so the mean will represent the proportion of TRUE values.

2..Character to numeric conversion leads to NAs because characters can't be interpreted as numbers.Because characters like "a", "b", etc. cannot be converted to numbers.

3..Factor to numeric conversion works by converting each factor level to its corresponding numeric code.                   